---
title: "Monte Carlo Simulation of Stock Portfolio in R, Matlab, and Python"
author: "Feichen Shen, Israel Diego, Shuoran Li"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    theme: journal
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Monte Carlo Introduction

The purpose of this tutorial is to demonstrate Monte Carlo Simulation in Matlab, R, and Python. We conduct our Monte Carlo study in the context of simulating daily returns 
for an investment portfolio.

For simplicity we will only consider three assets: Apple, Google, and Facebook. 
We will assume an Initial Investment of $100,000 and allocate our money evenly 
between the three stocks. In this case the portfolio weights $w_i = 1/3$ for 
$i = 1,2,3$. 

Next, we assume that daily returns are distributed Multivariate Normal with mean 
vector $\mu$ and covariance matrix $\Sigma$. In other words, 
$$R_t \sim MVN(\mu, \Sigma)$$
We will use the Cholesky Factorization in order to find Lower Triangular Matrix 
$L$ such that $LL' = \Sigma$. Then our returns can be generated by 
$$ R = \mu + LZ $$ where $$Z \sim N(0,I)$$.

The returns will be simulated over a 30-day period, where our 30-day returns
can be formulated as, $$R_{30} = \prod_{i=1}^{30} (1+R_i)$$

Thus our portfolio returns for each Monte Carlo trial $m$ become the inner 
product between the 30-day returns and our vector of portfolio weights $w$, 
$$P_m = w \cdot R_{30} $$.

## Dataset Summary

We use adjusted-close stock prices for Apple, Google, and Facebook from November 14th, 2017 - November 14th, 2018. Historical stock price data can be found on Yahoo 
Finance for these companies. Also [here](https://raw.githubusercontent.com/ShuoranLi/506_Project/master/Group21_ProjectData.csv) is the link to the data set for this
tutorial.

## {.tabset}

### R
```{r, code = readLines("FinalProject_R.R")}
```
(tab content)

### Matlab

(tab content)

### Python

(tab content)